# Managing multiple Rs on Mac

Reference: https://www.monicathieu.com/posts/2024-05-21-multiple-r-versions/

First, we need to ask Mac to forget the installed R's so that new ones can be installed without removing the old one.


Find all packages related to R:

```{}
pkgutil --pkgs='org.?-project*'
```

```{r, include=FALSE}
#system("pkgutil --pkgs='org.?-project*'")
```


```{}
org.R-project.arm64.R.GUI.pkg
org.r-project.R.el-capitan.fw.pkg
org.r-project.arm64.tcltk
org.r-project.x86_64.tcltk.x11
org.R-project.R.GUI.pkg
org.R-project.arm64.R.fw.pkg
org.r-project.x86_64.texinfo
org.r-project.arm64.texinfo
org.r-project.R.el-capitan.GUI.pkg
```

Forget them all, one by one, till noe shows up on rerun of the above code. Here is an example:

```{}
sudo pkgutil --forget org.R-project.arm64.R.GUI.pkg
```

Run the following to see the currently installed versions and the active one:

```{}
ls -lh /Library/Frameworks/R.framework/Versions
```

```{r, echo=FALSE, message=FALSE}
system('ls -lh /Library/Frameworks/R.framework/Versions', intern = TRUE) |> sapply(function(x) cat(x, '\n')) -> tmp
```

If you need to change the active one, use the following instructions:

```{}
# Symbolic Link: ln -s source destination
ln -s /Library/Frameworks/R.framework/Versions/4.1-arm64 /Library/Frameworks/R.framework/Versions/Current
```

## R, Miniconda and VSCode
Reference: https://www.codydehaan.com/blog/vscode-miniconda-r/

### Miniconda
Install. 

To disable auto activating `base` environment:
```{}
conda config --set auto_activate_base False
```

#### Cloning
References:

- https://stackoverflow.com/questions/40700039/how-can-you-clone-a-conda-environment-into-the-root-environment and 
- https://stackoverflow.com/questions/57701571/what-is-the-right-way-to-update-anaconda-and-conda-base-environments


To clone an environment `envA` as `envB` on the same device:
```{}
conda create --name envB --clone envA
```

To clone an `envA` as `envB` using a `yml` file:

```{}
conda activate envA
conda env export > enviroment_A.yml
conda env create -n envB --file environment_A.yml
```

Similarly, an text file can be used for closing:

```{}
conda activate envA
conda list --explicit > spec_file_A.txt
conda env create -n envB --file spec_file_A.txt
```

To update an environment, we can use `spec_file_A.txt` and `install` command:

```{}
conda install -n envB --file spec_file_A.txt
```


#### Updating
To update all environments:

```{}
conda update --all
```

To update a given environment:

```{}
conda update -n envB --all
```

To update an environment according to a `yml` file:

```{}
conda env update -n envB --file environment_A.yml  
```

#### Essential Packages

  - `r-base`
  - `radian`
  - `r-renv`
  - `r-httpgd`
  - `r-devtools`

Starting with R `4.2.0` I was able to install the above libraries via conda at the expense of upgrading to `4.3.1`. But the langugage server could not be installed.

  - `r-languageserver`
  - `r-tidyverse`

Packages installed by `conda` are not accessible within packages that are maintained by `renv`.
A work around is to install does packages from source file downloaded by `conda` (at `.cache`) so that they can be usedbut not tracked by `renv`.
#### `renv`

- `renv::init()`: To initiate `renv`.
  - Creates the `renv.lock` lockfile.
  - Creates/updates `.Rprofile`.
  - Creates `renv` folder with `renv/settings.json` and `renv/activate.R` files.
  
- `renv::snapshot()`: To update the `renv.lock` lockfile.
  - This will update the `renv.lock` particularly after new libraries are installed.
    - Using `renv::install()` reduces the number of downloads when multiple projects are managed on the same device.
    - Use `renv::update()` to update packages.

- `renv::restore()`: To recreate the same environment documented in `renv.lock`. It seems that the other three file also needs to be shared when cloning the environment.
  - Run this after each pull that includes changes to any of the `renv` files.
  

### VSCodium
Install `VSCodium` or `VS Code`.

Install `R Extension`.

Updating the extension settings:

- Enable Bracketed Paste
- Rpath: `Path-to-miniconda3/[envs/envName/]bin/R`
- Rterm: Mac — `Path-to-miniconda3/[envs/envName/]bin/radian`
- Rterm: Option — Add `--r-binary=Path-to-miniconda3/[envs/envName/]bin/R`
```

