# Managing multiple Rs on Mac

Reference: https://www.monicathieu.com/posts/2024-05-21-multiple-r-versions/

First, we need to ask Mac to forget the installed R's so that new ones can be installed without removing the old one.


Find all packages related to R:

```{}
pkgutil --pkgs='org.?-project*'
```

```{r, include=FALSE}
#system("pkgutil --pkgs='org.?-project*'")
```


```{}
org.R-project.arm64.R.GUI.pkg
org.r-project.R.el-capitan.fw.pkg
org.r-project.arm64.tcltk
org.r-project.x86_64.tcltk.x11
org.R-project.R.GUI.pkg
org.R-project.arm64.R.fw.pkg
org.r-project.x86_64.texinfo
org.r-project.arm64.texinfo
org.r-project.R.el-capitan.GUI.pkg
```

Forget them all, one by one, till noe shows up on rerun of the above code. Here is an example:

```{}
sudo pkgutil --forget org.R-project.arm64.R.GUI.pkg
```

Run the following to see the currently installed versions and the active one:

```{}
ls -lh /Library/Frameworks/R.framework/Versions
```

```{r, echo=FALSE, message=FALSE}
system('ls -lh /Library/Frameworks/R.framework/Versions', intern = TRUE) |> sapply(function(x) cat(x, '\n')) -> tmp
```

If you need to change the active one, use the following instructions:

```{}
# Symbolic Link: ln -s source destination
ln -s /Library/Frameworks/R.framework/Versions/4.1-arm64 /Library/Frameworks/R.framework/Versions/Current
```

