---
title: "Package Workflow"
author: "Mansour Abdoli, PhD"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
knitr::opts_chunk$set(eval = FALSE)
```

# File: setup_package.R

## 1. Initialize new R package
```{r}
library(usethis)
library(devtools)

```

### Create a new package (run once)
```{r}
create_package <- function(path, add2repo=TRUE) {
  usethis::create_package(path)
  usethis::use_git()       # Initialize git
  if(add2repo)
    usethis::use_github()    # Optional: Set up GitHub repository
  
  # Set up package infrastructure
  usethis::use_mit_license()    # Or another license
  usethis::use_readme_md()      # Create README.md
  usethis::use_news_md()        # Create NEWS.md for tracking changes
  usethis::use_package_doc()    # Package-level documentation
  usethis::use_testthat()       # Set up testing infrastructure
  #usethis::use_travis()         # Optional: Set up Travis CI
  #usethis::use_coverage()       # Optional: Set up test coverage
  
  # Create standard directories
  dir.create("inst", showWarnings = FALSE)
  dir.create("data", showWarnings = FALSE)
}
```
## 2. Set up .gitignore
```{r}
setup_gitignore <- function() {
  usethis::use_git_ignore(c(
    ".Rproj.user",
    ".Rhistory",
    ".RData",
    ".Ruserdata",
    "*.Rproj",
    ".DS_Store",    # For Mac users
    "inst/doc",
    "docs/",
    "*.tar.gz",
    "*.Rcheck/",
    "*.rdb",
    "*.rdx",
    "*.dll",
    "*.o",
    "*.so",
    "src/*.o",
    "src/*.so",
    "src/*.dll"
  ))
}
```

## 3. Set up Travis CI configuration
```{r}
setup_travis <- function() {
  writeLines(
    c(
      "language: r",
      "cache: packages",
      "",
      "r:",
      "  - release",
      "  - devel",
      "",
      "sudo: false",
      "",
      "after_success:",
      "  - Rscript -e 'covr::codecov()'",
      "",
      "notifications:",
      "  email:",
      "    on_success: change",
      "    on_failure: always"
    ),
    ".travis.yml"
  )
}
```


## 4. Create DESCRIPTION template
```{r}
setup_description <- function(pkg_name, 
                              Title="What the Package Does (One Line, Title Case)",
                              maintainer_name, maintainer_email,
                              Description = "What the package does (one paragraph).") {
  desc_template <- paste0(
    "Package: ", pkg_name, "\n",
    "Type: Package\n",
    "Title: ", Title,"\n",
    "Version: 0.1.0\n",
    "Authors@R: person('", maintainer_name, "',\n",
    "                  email = '", maintainer_email, "',\n",
    "                  role = c('cre', 'aut'))\n",
    "Description: ", Description, "\n",
    #"License: MIT + file LICENSE\n", # License will be added using usethis::xxx_license()
    "Encoding: UTF-8\n",
    "LazyData: true\n",
    "Roxygen: list(markdown = TRUE)\n",
    "RoxygenNote: 7.1.1\n",
    "Suggests:\n",
    "    testthat (>= 3.0.0),\n",
    "    knitr,\n",
    "    rmarkdown\n",
    "Config/testthat/edition: 3\n",
    "VignetteBuilder: knitr\n"
  )
  writeLines(desc_template, "DESCRIPTION")
}
```
## 5. Create development version script
```{r}
create_dev_version <- function() {
  usethis::use_version('dev')
}

# Example usage:
if (FALSE) {  # Prevent accidental execution
  # Initialize new package
  create_package("projUtils")
  setwd(paste0(getwd(), .Platform$file.sep ,"projUtils"))
  
  # Set up version control infrastructure
  setup_gitignore()
  setup_travis()
  setup_description("mypackage", "Your Name", "your.email@example.com")
  
  # Initial commit
  system("git add .")
  system('git commit -m "Initial commit"')
  
  # Create development version
  create_dev_version()
}
```


## 6. Example workflow functions
```{r}
increment_version <- function(type = c("patch", "minor", "major")) {
  type <- match.arg(type)
  usethis::use_version(type)
}

create_feature_branch <- function(feature_name) {
  system(paste0("git checkout -b feature/", feature_name))
}

merge_feature <- function(feature_name) {
  system("git checkout main")
  system(paste0("git merge feature/", feature_name))
}

prepare_release <- function(version) {
  # Update NEWS.md
  usethis::edit_file("NEWS.md")
  
  # Update version number
  desc::desc_set_version(version)
  
  # Run tests
  devtools::test()
  
  # Build documentation
  devtools::document()
  
  # Check package
  devtools::check()
}

create_release_tag <- function(version) {
  system(paste0('git tag -a v', version, ' -m "Release version ', version, '"'))
  system("git push --tags")
}
```

## 7. Example .Rbuildignore
```{r}
setup_rbuildignore <- function() {
  usethis::use_build_ignore(c(
    "^.*\\.Rproj$",
    "^\\.Rproj\\.user$",
    "^data-raw$",
    "^docs$",
    "^\\.travis\\.yml$",
    "^codecov\\.yml$",
    "^_pkgdown\\.yml$",
    "^README\\.Rmd$",
    "^cran-comments\\.md$",
    "^\\.github$"
  ))
}
```


# Handling Dependencies

## File: dependencies_management.R

```{r}
library(usethis)
library(desc)
library(devtools)
library(remotes)
```


## 1. Function to manage package dependencies
```{r}
manage_dependencies <- function(pkg_path = ".") {
  # Initialize description file if it doesn't exist
  if (!file.exists(file.path(pkg_path, "DESCRIPTION"))) {
    stop("DESCRIPTION file not found. Are you in a package directory?")
  }
  
  # Function to add a package dependency
  add_dependency <- function(pkg_name, type = c("Imports", "Suggests", "Depends", "Enhances", "LinkingTo")) {
    type <- match.arg(type)
    usethis::use_package(pkg_name, type = type)
  }
  
  # Function to add version constraints
  add_version_constraint <- function(pkg_name, version_constraint) {
    desc::desc_set_dep(pkg_name, version = version_constraint)
  }
  
  # Function to remove a package dependency
  remove_dependency <- function(pkg_name) {
    desc::desc_del_dep(pkg_name)
  }
  
  # Return functions in a list
  list(
    add = add_dependency,
    add_version = add_version_constraint,
    remove = remove_dependency
  )
}
```

## 2. Function to set up package namespace imports
```{r}
setup_namespace <- function() {
  # Create or update NAMESPACE file
  roxygen2::roxygenize()
  
  # Example namespace directives
  writeLines(c(
    "# Generated by roxygen2: do not edit by hand",
    "",
    "# Import specific functions",
    "#' @importFrom dplyr filter mutate select",
    "#' @importFrom ggplot2 ggplot geom_point theme_minimal",
    "",
    "# Import entire packages",
    "#' @import data.table",
    "",
    "# Export all functions with names that don't start with dot",
    "#' @exportPattern \"^[^\\\\.]\"",
    "",
    "# S3 method exports",
    "#' @export",
    "print.myclass <- function(x, ...) {",
    "  # method implementation",
    "}"
  ), "R/namespace_example.R")
}
```

## 3. Function to handle system dependencies
```{r}
check_system_dependencies <- function() {
  # Example system requirements check
  sys_reqs <- list(
    python = function() system("python --version") == 0,
    java = function() system("java -version") == 0,
    gcc = function() system("gcc --version") == 0
  )
  
  results <- lapply(names(sys_reqs), function(req) {
    result <- try(sys_reqs[[req]](), silent = TRUE)
    list(
      name = req,
      available = !inherits(result, "try-error") && result
    )
  })
  
  do.call(rbind, results)
}
```

## 4. Function to set up repository sources
```{r}
setup_repositories <- function() {
  # Add multiple repository sources
  options(
    repos = c(
      CRAN = "https://cloud.r-project.org",
      BioCsoft = "https://bioconductor.org/packages/release/bioc",
      BioCann = "https://bioconductor.org/packages/release/data/annotation",
      BioCexp = "https://bioconductor.org/packages/release/data/experiment"
    )
  )
}
```

## 5. Function to install development dependencies
```{r}
install_dev_dependencies <- function() {
  # Install common development tools
  devtools::install_deps(dependencies = TRUE)
  
  # Additional development packages
  dev_packages <- c(
    "testthat",
    "roxygen2",
    "covr",
    "devtools",
    "usethis",
    "pkgdown",
    "styler",
    "lintr"
  )
  
  installed <- rownames(installed.packages())
  to_install <- dev_packages[!dev_packages %in% installed]
  
  if (length(to_install) > 0) {
    install.packages(to_install)
  }
}
```


## 6. Function to create dependency documentation
```{r}
create_dependency_docs <- function() {
  # Read current dependencies
  desc <- desc::desc_get_deps()
  
  # Create markdown documentation
  doc <- c(
    "# Package Dependencies",
    "",
    "## Required Packages",
    "",
    "### Imports",
    paste("*", desc[desc$type == "Imports", "package"]),
    "",
    "### Suggests",
    paste("*", desc[desc$type == "Suggests", "package"]),
    "",
    "### System Requirements",
    "* R version >= 3.5.0",
    "* Python (optional, for Python integration)",
    "* gcc/g++ (for compilation of C++ code)",
    "",
    "## Installation",
    "```r",
    "# Install from CRAN",
    "install.packages('yourpackage')",
    "",
    "# Install development version",
    "remotes::install_github('username/yourpackage')",
    "```"
  )
  
  writeLines(doc, "dependencies.md")
}
```

## 7. Example usage for different dependency scenarios
```{r}
example_dependencies <- function() {
  deps <- manage_dependencies()
  
  # Add basic dependencies
  deps$add("dplyr", "Imports")      # Core dependency
  deps$add("ggplot2", "Imports")    # Visualization
  deps$add("testthat", "Suggests")  # Testing framework
  
  # Add version constraints
  deps$add_version("dplyr", ">= 1.0.0")
  deps$add_version("ggplot2", ">= 3.3.0")
  
  # Example of conditional dependency usage
  writeLines(c(
    "if (requireNamespace('ggplot2', quietly = TRUE)) {",
    "  # Code that uses ggplot2",
    "  plot <- ggplot2::ggplot(data, ggplot2::aes(x, y)) +",
    "    ggplot2::geom_point()",
    "} else {",
    "  # Fallback plotting code",
    "  plot <- plot(x, y)",
    "}"
  ), "R/conditional_example.R")
}
```

## 8. Function to check and update dependencies
```{r}
check_and_update_deps <- function() {
  # Check for outdated packages
  old <- old.packages()
  
  if (!is.null(old)) {
    cat("Outdated packages:\n")
    print(old[, c("Package", "Installed", "Available")])
    
    # Update packages
    update.packages(ask = FALSE)
  } else {
    cat("All packages are up to date.\n")
  }
}
```

# Example usage:
```{r}
if (FALSE) {  # Prevent accidental execution
  # Initialize dependency management
  deps <- manage_dependencies()
  
  # Add core dependencies
  deps$add("dplyr", "Imports")
  deps$add("ggplot2", "Imports")
  
  # Add development dependencies
  install_dev_dependencies()
  
  # Set up repositories
  setup_repositories()
  
  # Check system dependencies
  check_system_dependencies()
  
  # Create documentation
  create_dependency_docs()
  
  # Check for updates
  check_and_update_deps()
  
  #
  devtools::document(roclets = c('rd', 'collate', 'namespace'))
  
  # Clean and Build
  system("cd ..")
  system("R CMD INSTALL --preclean --no-multiarch --with-keep.source projUtils")
  system("cd projUtils")
  
  
}
```
